A Survey of Exoskeleton Control Architectures and Algorithms\\ (Draft 3.0)

### 2015-09-29 exos/hal.tex
more edits

keywords: model-based control; predefined gait trajectory; strength augmentation; rehabilitation.\\

@@

Compared to sensitivity amplification methods, HAL's use of EMG data allows it to estimate intention without amplifying external disturbances.  While direct sensing of user intention is ideal for combat scenarios, sEMG data is extremely difficult to work with due to high filtering and calibration requirements.  A hybrid strategy which uses sensitivity amplification and sEMG data (possibly thresholded) to filter user intention could prove effective.  As an alternative, research into nano-fabrication and MEMs technologies has produced new ``robotic skin'' sensing arrays. This technology may allow direct measurement of interaction forces (i.e. estimating pressure / force and exoskeleton contact locations), which has not been feasible in the past.  Such measurements could (potentially) filter external disturbances from user generated input to estimate intent.

>>>>

% keywords: model-based control; predefined gait trajectory; strength augmentation; rehabilitation.\\

@@

Compared to sensitivity amplification methods, HAL's use of EMG data allows it to estimate intention without amplifying external disturbances.  While direct sensing of user intention is ideal for noisy, contract-rich environments, sEMG data is extremely difficult to work with due to high filtering and calibration requirements.  A hybrid strategy which uses sensitivity amplification and sEMG data (possibly thresholded) to filter user intention could prove effective.  Additionally, Section~\ref{survey:recommend} mentions new capabilities in nano-fabrication and MEMs technologies that have produced new high-density, wearable sensing arrays. These and similarly advanced sensing technology may facilitate direct measurement of interaction forces (i.e. estimating pressure / force and exoskeleton contact locations), which could potentially filter external disturbances from user generated input to estimate intent.

### 2015-09-29 exos/xor.tex

In implementation, the XoR uses rubber air muscles (MDSP-40) connected to pulleys via tendons and to EP regulators in the backpack (a constant pressure and cam system is being tested to reduce weight required for backpack air valves).  The electric actuator is a geared, 200W brushless DC motor (Maxon EC powermax 30) rated at 4.7 A, which transmits power to joints through a belt and pulley system with a 57.5 gear ratio. The motor is backdrivable with 0.12 Nm output torque and up to 34.5 Nm (for short duration) at every joint.  The design yields a relatively high range of motion (up to 120 degrees) with a 150 Nm load capacity.  However, experiments reveal the need for bi-directional pneumatic actuation at hip joints.

For sensing, XoR is equipped with \textbf{rotary encoders} at joints, an \textbf{IMU} in the backpack, \textbf{load cells} in the feet, and is networked to servers capable of providing \textbf{EMG}, NIRS, and EEG data.  A control PC running at 1 KHz, pulse counter, amplifiers, Digital IO are external and so not included in the weight of the unit.  Human operators are outfitted with a \textbf{goniometer}.

@@

The XoR prototype uses a promising hybrid actuation scheme to achieve both fast peak torque response and higher sustained torque while remaining lightweight.  It tracks human motion with feedback from a goniometer and feedforward control provided by EMG data.  These human intention / sensing modalities are both difficult to implement due to noise and calibration issues.  However, capturing the user intent is important, especially in the presence of external disturbances.  If the sensory system were improved (e.g, wearable ``robot skin'' sensing arrays), this exoskeleton design could be highly effective.

>>>>

In implementation, the XoR uses rubber air muscles (MDSP-40) connected to pulleys via tendons and to electro-pneumatic regulators in the backpack (a constant pressure and cam system is being tested to reduce weight required for backpack air valves).  The electric actuator is a geared, 200W brushless DC motor (Maxon EC powermax 30) rated at 4.7 A, which transmits power to joints through a belt and pulley system with a 57.5 gear ratio. The motor is backdrivable with 0.12 Nm output torque and up to 34.5 Nm (for short duration) at every joint.  The design yields a relatively high range of motion (up to 120 degrees) with a 150 Nm load capacity.  However, experiments reveal the need for bi-directional pneumatic actuation at hip joints.

For sensing, XoR is equipped with \textbf{rotary encoders} at joints, an \textbf{IMU} in the backpack, \textbf{load cells} in the feet, and is networked to servers capable of providing \textbf{EMG}, NIRS (near-infrared spectroscopy), and EEG (electroencephalogram) data.  A control PC running at 1 KHz, pulse counter, amplifiers, Digital IO are external and so not included in the weight of the unit.  Human operators are outfitted with a \textbf{goniometer}.

@@

The XoR prototype uses a promising hybrid actuation scheme to achieve both fast peak torque response and higher sustained torque while remaining lightweight.  It tracks human motion with feedback from a goniometer and feedforward control provided by EMG data.  These human intention / sensing modalities are both difficult to implement due to noise and calibration issues.  However, capturing the user intent is important, especially in the presence of external disturbances.  If the sensory system were improved (see the discussion in Section~\ref{survey:recommend}), this exoskeleton design could be highly effective.

### 2015-09-29 survey.tex

\section{Conclusions and Recommendations}

All the (strength augmentation) exoskeletons we reviewed can be categorized into one of two general control categories.  The exoskeleton either directly measures user intent (e.g., from EMG data or human-exoskeleton interaction forces), or it estimates intent from measurements at exoskeleton joints and applies a sensitivity amplification strategy to shadow the wearer.  The first strategy is difficult to implement using standard sensing modalities.  For instance, EMG data is notoriously noisy and presents both modeling and calibration challenges.  It would also be extremely difficult to keep these sensors in place to obtain accurate readings in a combat scenario.  Similarly, many avoid the use of load cells for measurement of the human-exoskeleton interface fores to avoid outfitting the complicated sensor network required to accurate obtain interaction forces / torques (humans are compliant thus rigid attachment assumptions are typically inadequate).

While sensitivity amplification strategies avoid these sensing issues by estimating human intent from exoskeleton joint readings, they cannot distinguish between user generated torques and those produced by the environment.  As such, they end up amplifying noise and the user must react to stabilize the system.  This is also unacceptable for a combat exoskeleton.

@@

% \input{exos/exo-ul7}

%\section{Discussion}
%
%Discussion to be written.
%
%\section{Conclusions and Recommendations}
%
%All the (strength augmentation) exoskeletons we reviewed can be categorized into one of two general control categories.  The exoskeleton either directly measures user intent (e.g., from EMG data or human-exoskeleton interaction forces), or it estimates intent from measurements at exoskeleton joints and applies a sensitivity amplification strategy to shadow the wearer.  The first strategy is difficult to implement using standard sensing modalities.  For instance, EMG data is notoriously noisy and presents both modeling and calibration challenges.  It would also be extremely difficult to keep these sensors in place to obtain accurate readings in a combat scenario.  Similarly, many avoid the use of load cells for measurement of the human-exoskeleton interface fores to avoid outfitting the complicated sensor network required to accurate obtain interaction forces / torques (humans are compliant thus rigid attachment assumptions are typically inadequate).
%
%While sensitivity amplification strategies avoid these sensing issues by estimating human intent from exoskeleton joint readings, they cannot distinguish between user generated torques and those produced by the environment.  As such, they end up amplifying noise and the user must react to stabilize the system.  This is also unacceptable for a combat exoskeleton.
%

% \bibliographystyle{plain}
% \bibliography{exo}

%\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

>>>>

\section{Conclusions and Recommendations}
\label{survey:recommend}

All the (strength augmentation) exoskeletons we reviewed can be categorized into two general control categories.  The exoskeleton either directly measures user intent (e.g., from EMG data or human-exoskeleton interaction forces), or it estimates intent from measurements at exoskeleton joints and applies a sensitivity amplification strategy to shadow the wearer.  The first strategy is difficult to implement using standard sensing modalities.  For instance, EMG data is notoriously noisy and presents both modeling and calibration challenges.  It would also be extremely difficult to keep these sensors in place to obtain accurate readings in dynamic environments or combat scenarios.  Similarly, many avoid the use of load cells for measurement of human-exoskeleton interface forces and the complicated sensor network required to accurately obtain interaction forces / torques (humans are compliant thus rigid attachment assumptions are typically inadequate).

While sensitivity amplification strategies circumvent these sensing issues by estimating human intent from exoskeleton joint readings, they cannot distinguish between user generated torques and those produced by the environment.  As such, they end up amplifying noise and the user must react to stabilize the system.  This is also unacceptable for highly dynamic exoskeletons operating in contact-rich environments.

We emphasize that user intent is critical, and current sensing challenges must be addressed in order to develop dynamic exoskeletons capable of serving in combat scenarios.  As a stop-gap, it may be possible to combine low information density EMG (thresholded on/off muscle activation signals) and human force/torque or angular measurements to filter user intent.  Such an approach may prove useful in a hybrid strategy, filtering external disturbance forces during sensitivity amplification control.  As a longer term solution, we highlight research in the area of MEMs and nano-fabrication, which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability. We recommend further study and the development of more advanced sensory systems to address the challenges inherent in dynamic exoskeleton control.

In addition to capturing user intent, it is important to develop accurate models of the human/exoskeleton system.  Feedback strategies, especially those based on force/torque control loops and series elastic actuation, are limited in bandwidth. Highly dynamic behaviors require feedforward models to reduce reliance on feedback and to allow lower-impedance feedback, which reduces resistive forces acting on the user.

%% OTHER POINTS:
% Model-based approaches are heavily reliant on accurate models 
% Underactuated -- feedback linearization typically assumes fully actuated systems.

@@
